|

### 2021-11-23 01-Chp-Background.Rmd
End of day:
Citations fixed!

@@

Defined by the Diagnostic and Statistical Manual of Mental Disorders as a "major neurocognitive disorder", dementia is a progressive disease which impairs cognitive functions including speech, memory and executive reasoning.[@edition2013diagnostic] At advanced stage, the condition causes severe behavioral and personality changes,[@cerejeira2012] cumulating in reduced motor control that affects patients ability to swallow or breathe.[@kumar2013] The condition has several distinct underlying causes, including Alzheimer's disease and vascular dementia.[@burns2009]

@@

<!----------------------------------------------------------------------------->
(ref:diagnosticCriteria-caption) Overview of the DSM-5 criteria for dementia and vascular dementia.[@edition2013diagnostic]

@@

One of the most commonly used criteria are those found in the Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria (Table \@ref(tab:diagnosticCriteria-table)).[@edition2013diagnostic] These criteria are outlined in Table \@ref(tab:diagnosticCriteria-table), and form the broad definition of a dementia diagnoses, supported by a detailed patient history, evidence from carers and family members, and objective assessments of cognitive ability using neurocognitive tests.

@@

Differentiating between the underlying causes of a dementia diagnosis is challenging but necessary, as whether the patient has Alzheimer's disease or vascular dementia will affect expected progression and potential treatment options available (see Section \@ref(intro-treatments)). Cause-specific criteria exist for the diagnosis of dementia sub-types. For example, the NINCDS-ADRDA criteria are commonly used to assess patients for Alzheimer's disease,[@dubois2007] while vascular dementia is diagnosed using the NINCDS-AIREN criteria.[@roman1993vascular]

@@

Dementia also has a substantial economic impact. In 2015, the estimated total cost of dementia in England was £24.2 billion. Health care costs alone were £3.8 billion,[@wittenberg2019] Globally, the cost of dementia care is expected to rise to $1tr by 2030.[@prince2014dementia]

@@

Given the substantial burden that dementia represents and the absence of any curative therapies, as detailed in the above sections, the assessment of easily modifiable targets for their utility in the prevention of dementia should be prioritized.[@winblad2016a] To date, a substantial amount of research has been produced examining putative risk factors for dementia.[@feingold2000;@peters2019; @anstey2019]

The benefits of a prevention-based approach based on addressing these risk factors are well-studied. Reducing the prevalence of the seven most important risk factors for dementia (obesity, hypertension,[@hughes2020association] diabetes, smoking, physical inactivity, and low educational attainment) by 10-20% per decade is estimated to result in a reduction in dementia prevalence of 8-15% by 2050.[@norton2014potential]

In this context, lipid levels represent a promising target for preventative treatment, due to the ready availability of lipid-modifying treatments which could be repurposed.[@pushpakom2019a] Determining whether variations in lipid levels are causative for dementia may prove critical in reducing the future burden of the condition.

@@

Several statin treatments have been widely available for some time (see Table \@ref(tab:statinOverview-table)). Depending on the statin and dosage prescribed, the average reduction in LDL-c concentrations ranges from 15% with low-intensity regimen (e.g. ravastatin 5 mg/day) up to 60% with a high-intensity regimen (e.g. rosuvastatin 80 mg/day).[@collins2016a; @law2003] Statins also vary with regard to their lipophilicity (the extent to which they are lipid soluble), affecting their localisation within the body, with hydophilic statins being concentrated in the liver and lipophilic statins circulating more widely.[@schachter2005] This may create a divide in the pleiotropic affects of statins with differing lipophilicity, particularly given the ability of lipophilic statins to permeate the blood brain barrier.[@sierra2011]

@@

Other therapies targeting triglycerides exist, including nicotinic acids[@mckenney2004new] and omega-3-fatty acids,[@skulas-rayannc.2019] but they far less effective in LDL-c lowering than the therapies described above.

@@

By far the largest source of evidence on the relationship between comes from observational designs. Several studies have examined the relationships between concentrations of serum lipids (total cholesterol (TC), low density lipoprotein cholesterol (LDL-c), high density lipoprotein cholesterol (HDL-c) and triglycerides) and both Alzheimer's disease and vascular dementia and reported extremely varied results. In some studies, a high serum cholesterol concentration has been found to be associated with an increase in susceptibility to AD,[@kivipelto2002; @kivipelto2005; @schilling2017; @solomon2009; @whitmer2005] however others have shown no association,[@li2005a; @mainous2005; @mielke2010; @tan2003a] or a reduced susceptibility [@mielke2005; @reitz2004a]. With regards vascular dementia, decreased levels of HDL-c appear to be associated with increased risk [@reitz2004a], while for LDL-c, studies have reported both positive and negative associations.[@reitz2004a; @moroney1999]

Several previous systematic review of observational studies examining the effect of lipids[@anstey2015] and lipid-regulating agents[@chu2018;@poly2020c] on dementia outcomes have been performed. However, these reviews have several limitations. Many did not consider grey literature sources (see Section \@ref(diverse-sources-preprints)). Additionally, many of the reviews of observational studies did not perform any risk-of-bias assessment[@chu2018b] or used an outdated assessment tool.[@anstey2015;@poly2020c]

@@

However, RCTs may be infeasible if the outcome of interest is one with a long prodomal period, such as dementia (see Section \@ref(underlying-pathologies)), as they would require extremely long and costly follow-up.[@ritchie2015] It is no surprise then that the two previous trials providing evidence on the effect of statins on dementia risk, identified by a recent Cochrane review,[@mcguinness2016a] are in fact trials of statins for the prevention of coronary related outcomes.

While being widely cited, these studies have major limitations that reduce their utility as a source of evidence on the effect of statin treatment on in assessing the impact of lipid-lowering treatment on dementia risk. Firstly, there was no clinical cognitive evaluation of patients to determine a dementia outcome. One of the trials, the Prospective Study of Pravastatin in the Elderly (PROSPER) trial,[@trompet2010] reported not on dementia outcomes but on the change in cognitive scores over a mean of 3.2 years. As highlighted in Section \@ref(diagnostic-criteria), a "change in score" alone is insufficient to diagnose a dementia outcome. The second trial, the Medical Research Council/British Health Foundation Protection Study,[@2002] found no effect of simvastatin on dementia (`r estimate(1, 0.61, 1.65)`), but did not report how the outcome was assessed/recorded within the trial.

@@

Both trials were also limited by the relatively short follow-up period examined, expected when the primary outcome of the trials were coronary related conditions rather than dementia.[@2002; @trompet2010] The PROSPER trial had a mean follow-up of 3.2 years, while the MRC/BHF Protection Study estimated risk at 5 years of follow-up. Given the long lag time between non-symptomatic onset of dementia and clinical presentation, it is likely that these durations are insufficient to fully capture the onset of dementia. Finally, as they included only patients at high vascular risk, their generalisability to other settings is limited.[@mcguinness2016b]

@@

Recent MR studies indicated that genetically determined low levels of LDL-c may cause a reduction in AD risk.[@larsson2017c; @ostergaard2015] However, the effect was attenuated in sensitivity analysis that exclude the region surrounding the ApoE gene, the strongest known risk factor for Alzheimer's disease.[@kim2009] Inclusion of ApoE4 variants invalidates the exclusion restriction criteria (Assumption 3, above), as the risk reduction observed may be driven by variants in this region via a pathway independent of lipid levels. This was supported by further MR studies where _ApoE4_ variants were intentionally excluded.[@benn2017]

@@

Evidence synthesis is the process of finding and integrating information from several sources to examine a research question.[@donnelly2018a] A common tyoe of evidence synthesis is a systematic review, either with or without a meta-analysis.[@chandler2019chapter]

@@

A important developing source of grey literature are preprints. Defined by the Committee on Publication Ethics (COPE) as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process'[@committeeonpublicationethicscope2018], preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months,[@vale2016] to rapidly disseminate research findings, as occurred during the COVID-19 pandemic,[@fraser2020a] and to make available publications that may not have been accepted elsewhere in an attempt to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

One of the major criticisms of using preprints as an evidence source is that they have not yet undergone formal peer review.[@maslove2018; @schalkwyk2020] However, this approach assigns substantial weight to peer-review as a indicator of "quality", and is at odds with the acceptance of non-reviewed conference proceedings as an evidence source.[@lefebvre2019searching; @mahood2014] The argument for including preprints as an evidence source is further strengthened by results that demonstrate preprinted studies seldom change following peer review. Meta-studies of the concordance between preprinted and published studies showed that results were broadly comparable between the two, indicating that while the numerical results may change, the overall interpretation of the results were consistent in the majority of cases.[@shi2021a; @klein2019; @nicholson2021] This indicates that preprints should be considered a reliable reflection of a given study.

@@

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. However, each method is limited by its own biases. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these inherent differences in study design, and as a result, in biases.[@lawlor2016a] If several sources of evidence are available and point towards identical conclusions about an exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, strengthen the effect of the exposure and the other to attenuate it.[@lawlor2016a] As such, triangulating these results can provides us a middle-ground between the competing directions of bias. A triangulation approach can also prove useful in a prospective manner, helping to design new studies that are at risk of different sources of bias to that already available from the published literature.[@munafo2018]

@@

All existing evidence, regardless of study design, is first identified by the by the systematic review presented inChapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Risk-of-bias assessment using a domain-based tool is already a recommended part of the systematic review process, but is particularly important to a triangulation exercise. [@Pagen71; @sterne2019a; @mcguinness2018] As such, a core component of the review is a comprehensive domain-based risk-of-bias assessment for all included studies.

@@

Despite their advantages, IPD meta-analysis are rarely performed.[@tugwell2010] Factors limiting their uptake include the increased time and effort they require when compared to a summary-level analysis, and the low success rate associated with obtaining the raw data.[@nevitt2017a; @ventresca2020] The data underlying primary studies are frequently not publicly available,[@alsheikh-ali2011; @federer2018a] and the availability of data "available on request from authors" declines rapidly over time.[@vines2014] Several systematic barriers to open data sharing have been identified[@vanpanhuis2014a]. Of particular concern for biomedical IPD analyses are legal issues surrounding the sharing of medical data, motivated by concerns around patient privacy.[@wartenberg2010]

>>>>

<!-- TODO Contributor statements -->

@@

Defined by the Diagnostic and Statistical Manual of Mental Disorders as a "major neurocognitive disorder", dementia is a progressive disease which impairs cognitive functions including speech, memory and executive reasoning.[@edition2013] At advanced stage, the condition causes severe behavioral and personality changes,[@cerejeira2012] cumulating in reduced motor control that affects patients ability to swallow or breathe.[@kumar2013] The condition has several distinct underlying causes, including Alzheimer's disease and vascular dementia.[@burns2009]

@@

<!----------------------------------------------------------------------------->
(ref:diagnosticCriteria-caption) Overview of the DSM-5 criteria for dementia and vascular dementia.[@edition2013]

@@

One of the most commonly used criteria are those found in the Diagnostic and Statistical Manual of Mental Disorders (DSM) criteria (Table \@ref(tab:diagnosticCriteria-table)).[@edition2013] These criteria are outlined in Table \@ref(tab:diagnosticCriteria-table), and form the broad definition of a dementia diagnoses, supported by a detailed patient history, evidence from carers and family members, and objective assessments of cognitive ability using neurocognitive tests.

@@

Differentiating between the underlying causes of a dementia diagnosis is challenging but necessary, as whether the patient has Alzheimer's disease or vascular dementia will affect expected progression and potential treatment options available (see Section \@ref(intro-treatments)). Cause-specific criteria exist for the diagnosis of dementia sub-types. For example, the NINCDS-ADRDA criteria are commonly used to assess patients for Alzheimer's disease,[@dubois2007] while vascular dementia is diagnosed using the NINCDS-AIREN criteria.[@roman1993]

@@

Dementia also has a substantial economic impact. In 2015, the estimated total cost of dementia in England was £24.2 billion. Health care costs alone were £3.8 billion,[@wittenberg2019] Globally, the cost of dementia care is expected to rise to $1tr by 2030.[@prince2014]

@@

Given the substantial burden that dementia represents and the absence of any curative therapies, as detailed in the above sections, the assessment of easily modifiable targets for their utility in the prevention of dementia should be prioritized.[@winblad2016] To date, a substantial amount of research has been produced examining putative risk factors for dementia.[@feingold2000;@peters2019; @anstey2019]

The benefits of a prevention-based approach based on addressing these risk factors are well-studied. Reducing the prevalence of the seven most important risk factors for dementia (obesity, hypertension,[@hughes2020] diabetes, smoking, physical inactivity, and low educational attainment) by 10-20% per decade is estimated to result in a reduction in dementia prevalence of 8-15% by 2050.[@norton2014potential]

In this context, lipid levels represent a promising target for preventative treatment, due to the ready availability of lipid-modifying treatments which could be repurposed.[@pushpakom2019] Determining whether variations in lipid levels are causative for dementia may prove critical in reducing the future burden of the condition.

@@

Several statin treatments have been widely available for some time (see Table \@ref(tab:statinOverview-table)). Depending on the statin and dosage prescribed, the average reduction in LDL-c concentrations ranges from 15% with low-intensity regimen (e.g. ravastatin 5 mg/day) up to 60% with a high-intensity regimen (e.g. rosuvastatin 80 mg/day).[@collins2016; @law2003] Statins also vary with regard to their lipophilicity (the extent to which they are lipid soluble), affecting their localisation within the body, with hydophilic statins being concentrated in the liver and lipophilic statins circulating more widely.[@schachter2005] This may create a divide in the pleiotropic affects of statins with differing lipophilicity, particularly given the ability of lipophilic statins to permeate the blood brain barrier.[@sierra2011]

@@

Other therapies targeting triglycerides exist, including nicotinic acids[@mckenney2004] and omega-3-fatty acids,[@skulas2019] but they far less effective in LDL-c lowering than the therapies described above.

@@

By far the largest source of evidence on the relationship between comes from observational designs. Several studies have examined the relationships between concentrations of serum lipids (total cholesterol (TC), low density lipoprotein cholesterol (LDL-c), high density lipoprotein cholesterol (HDL-c) and triglycerides) and both Alzheimer's disease and vascular dementia and reported extremely varied results. In some studies, a high serum cholesterol concentration has been found to be associated with an increase in susceptibility to AD,[@kivipelto2002; @kivipelto2005; @schilling2017; @solomon2009; @whitmer2005] however others have shown no association,[@li2005; @mainous2005; @mielke2010; @tan2003] or a reduced susceptibility [@mielke2005; @reitz2004]. With regards vascular dementia, decreased levels of HDL-c appear to be associated with increased risk [@reitz2004], while for LDL-c, studies have reported both positive and negative associations.[@reitz2004; @moroney1999]

<!-- Several previous systematic review of observational studies examining the effect of lipids[@anstey2015] and lipid-regulating agents[@chu2018;@poly2020] on dementia outcomes have been performed. However, these reviews have several limitations. Many did not consider grey literature sources (see Section \@ref(diverse-sources-preprints)). Additionally, many of the reviews of observational studies did not perform any risk-of-bias assessment[@chu2018] or used an outdated assessment tool.[@anstey2015;@poly2020c] -->

@@

However, RCTs may be infeasible if the outcome of interest is one with a long prodomal period, such as dementia (see Section \@ref(underlying-pathologies)), as they would require extremely long and costly follow-up.[@ritchie2015] It is no surprise then that the two previous trials providing evidence on the effect of statins on dementia risk, identified by a recent Cochrane review,[@mcguinness2016] are in fact trials of statins for the prevention of coronary related outcomes.

While being widely cited, these studies have major limitations that reduce their utility as a source of evidence on the effect of statin treatment on in assessing the impact of lipid-lowering treatment on dementia risk. Firstly, there was no clinical cognitive evaluation of patients to determine a dementia outcome. One of the trials, the Prospective Study of Pravastatin in the Elderly (PROSPER) trial,[@trompet2010] reported not on dementia outcomes but on the change in cognitive scores over a mean of 3.2 years. As highlighted in Section \@ref(diagnostic-criteria), a "change in score" alone is insufficient to diagnose a dementia outcome. The second trial, the Medical Research Council/British Health Foundation Protection Study,[@heartprotectionstudycollaborativegroup2002] found no effect of simvastatin on dementia (`r estimate(1, 0.61, 1.65)`), but did not report how the outcome was assessed/recorded within the trial.

@@

Both trials were also limited by the relatively short follow-up period examined, expected when the primary outcome of the trials were coronary related conditions rather than dementia.[@heartprotectionstudycollaborativegroup2002; @trompet2010] The PROSPER trial had a mean follow-up of 3.2 years, while the MRC/BHF Protection Study estimated risk at 5 years of follow-up. Given the long lag time between non-symptomatic onset of dementia and clinical presentation, it is likely that these durations are insufficient to fully capture the onset of dementia. Finally, as they included only patients at high vascular risk, their generalisability to other settings is limited.[@mcguinness2016]

@@

Recent MR studies indicated that genetically determined low levels of LDL-c may cause a reduction in AD risk.[@larsson2017; @ostergaard2015] However, the effect was attenuated in sensitivity analysis that exclude the region surrounding the ApoE gene, the strongest known risk factor for Alzheimer's disease.[@kim2009] Inclusion of ApoE4 variants invalidates the exclusion restriction criteria (Assumption 3, above), as the risk reduction observed may be driven by variants in this region via a pathway independent of lipid levels. This was supported by further MR studies where _ApoE4_ variants were intentionally excluded.[@benn2017]

@@

Evidence synthesis is the process of finding and integrating information from several sources to examine a research question.[@donnelly2018] A common tyoe of evidence synthesis is a systematic review, either with or without a meta-analysis.[@chandler2019chapter]

@@

A important developing source of grey literature are preprints. Defined by the Committee on Publication Ethics (COPE) as 'scholarly manuscript[s] posted by the author(s) in an openly accessible platform, usually before or in parallel with the peer review process'[@committeeonpublicationethicscope2018], preprints serve several purposes. They are used to establish primacy when submitting to a journal where the peer-review process may take several months,[@vale2016] to rapidly disseminate research findings, as occurred during the COVID-19 pandemic,[@fraser2020preprinting] and to make available publications that may not have been accepted elsewhere in an attempt to combat publication bias or the "file-drawer" effect.[@rosenthal1979]

One of the major criticisms of using preprints as an evidence source is that they have not yet undergone formal peer review.[@maslove2018; @schalkwyk2020] However, this approach assigns substantial weight to peer-review as a indicator of "quality", and is at odds with the acceptance of non-reviewed conference proceedings as an evidence source.[@lefebvre2019searching; @mahood2014] The argument for including preprints as an evidence source is further strengthened by results that demonstrate preprinted studies seldom change following peer review. Meta-studies of the concordance between preprinted and published studies showed that results were broadly comparable between the two, indicating that while the numerical results may change, the overall interpretation of the results were consistent in the majority of cases.[@shi2021; @klein2019; @nicholson2021] This indicates that preprints should be considered a reliable reflection of a given study.

@@

As illustrated in Section \@ref(evidence-association), several diverse epidemiological methods have been used to examine the effect of varying blood lipid levels on dementia risk. However, each method is limited by its own biases. Aetiological triangulation is a developing evidence synthesis method that seeks to exploit these inherent differences in study design, and as a result, in biases.[@lawlor2016] If several sources of evidence are available and point towards identical conclusions about an exposure-outcome relationship, and these sources are at risk of unrelated biases, this strengthens our confidence in the result. The ideal scenario is where predicted sources of bias are likely to be in competing directions, strengthen the effect of the exposure and the other to attenuate it.[@lawlor2016] As such, triangulating these results can provides us a middle-ground between the competing directions of bias. A triangulation approach can also prove useful in a prospective manner, helping to design new studies that are at risk of different sources of bias to that already available from the published literature.[@munafo2018]

@@

All existing evidence, regardless of study design, is first identified by the by the systematic review presented inChapters \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading). Risk-of-bias assessment using a domain-based tool is already a recommended part of the systematic review process, but is particularly important to a triangulation exercise. [@Pagen71; @sterne2019; @mcguinness2018] As such, a core component of the review is a comprehensive domain-based risk-of-bias assessment for all included studies.

@@

Despite their advantages, IPD meta-analysis are rarely performed.[@tugwell2010] Factors limiting their uptake include the increased time and effort they require when compared to a summary-level analysis, and the low success rate associated with obtaining the raw data.[@nevitt2017; @ventresca2020] The data underlying primary studies are frequently not publicly available,[@alsheikh-ali2011; @federer2018] and the availability of data "available on request from authors" declines rapidly over time.[@vines2014] Several systematic barriers to open data sharing have been identified[@vanpanhuis2014a]. Of particular concern for biomedical IPD analyses are legal issues surrounding the sharing of medical data, motivated by concerns around patient privacy.[@wartenberg2010]

### 2021-11-23 02-Chp-Systematic-Review-Tools.Rmd
End of day:
Citations fixed!

Searching these preprints as part of the systematic review described in Chapter \@ref(sys-rev-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the often complex Boolean logic (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018a;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search were to go through each record, one-by-one, downloading individual citations. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

The tool has been well received by the community (as of `r date`, `medrxivr` has been downloaded more than `r medrxivr_downloads` times), and several use cases have been reported. It has been used to investigate the role of preprints in the response to the 2019 coronavirus outbreak,[@kodvanj2020] perform searches of preprints as part of a systematic review,[@noone2020; @grassly2020] and examine how data-sharing behaviour is affected by journal policies (see \@ref(thesis-output)).[@mcguinness2020c]

The package has been accepted into the rOpenSci suite of packages, a collection of "carefully vetted, staff- and community-contributed R software tools that lower barriers to working with scientific data sources on the web".[@boettiger2015] As part of this process, following rigorous peer-review, an associated article introducing the tool was published by the Journal of Open Source Software.[@mcguinness2020a] The entire review discussion is publicly available and can be viewed online.[@zotero-15016] The tool has also been well received by the open-source community, demonstrated by the engagement of other developers in contributing to important new functionality and suggesting bug-fixes.

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-sys-rev-tools)). `r publish` of the `r publication_n` records posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting,[@abdill2019b] indicating that a non-insignificant number of preprints are never formally published but remain accessible as preprints.

@@

@@

Part of the motivation for creating the `medrxivr` tool was a belief that the development and distribution of open source scripts and tools should be a fundamental part of evidence synthesis research.[@goldacre2019b; @mckiernan2016c] In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts that have a similar, or related, functionality - in fact, following development of the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package, which allows for importing medRxiv metadata into R but does not provide search functionality).[@rbiorxiv] If these scripts continue to be developed in private and are never shared or publicised, this will inevitably hamper the efforts of evidence synthesis community, not only in terms of duplication of time and effort but also due to lost opportunities for collaboration.[@mckiernan2016c] Creating and sharing well-documented packages, the recognised standard for sharing code in R, represents one way to reduce this inefficiency.[@vuorre2020]

>>>>

Searching these preprints as part of the systematic review described in Chapter \@ref(sys-rev-heading) was a necessity, as many of the existing reviews on the topic of lipids and dementia have not considered this important source of evidence. At the time of writing, however, the bioRxiv and medRxiv websites allow only simple search queries as opposed to the often complex Boolean logic (AND/OR/NOT) that information specialists use to query other major databases.[@bramer2018;@gusenbauer2020] Additionally, the best available extraction mechanism for obtaining references for all records returned by a search were to go through each record, one-by-one, downloading individual citations. As the scale of these preprint databases increase, particularly in light of the massive expansion of the medRxiv repository as a result of COVID, this already time-consuming and error-prone method is no longer feasible.

@@

The tool has been well received by the community (as of `r date`, `medrxivr` has been downloaded more than `r medrxivr_downloads` times), and several use cases have been reported. It has been used to investigate the role of preprints in the response to the 2019 coronavirus outbreak,[@kodvanj2020] perform searches of preprints as part of a systematic review,[@noone2020; @grassly2020] and examine how data-sharing behaviour is affected by journal policies (see \@ref(thesis-output)).[@mcguinness2020DAScomparison]

The package has been accepted into the rOpenSci suite of packages, a collection of "carefully vetted, staff- and community-contributed R software tools that lower barriers to working with scientific data sources on the web".[@boettiger2015] As part of this process, following rigorous peer-review, an associated article introducing the tool was published by the Journal of Open Source Software.[@mcguinness2020medrxivr] The entire review discussion is publicly available and can be viewed online.[@zotero-15016] The tool has also been well received by the open-source community, demonstrated by the engagement of other developers in contributing to important new functionality and suggesting bug-fixes.

@@

Secondly, using `medrxivr`, an analysis of the publication rate for medRxiv preprints was performed (see Appendix \@ref(appendix-sys-rev-tools)). `r publish` of the `r publication_n` records posted on medRxiv in July 2019 were published by 30th July 2021 (i.e. allowing for a two-year lag between preprint posting and publication). This finding agrees with previous work demonstrating that two-thirds of bioRxiv preprints are published in a peer-reviewed journal within two years of posting,[@abdill2019popularity] indicating that a non-insignificant number of preprints are never formally published but remain accessible as preprints.

@@

<!-- TODO EMBASE will index medRxiv/bioRxiv from Christmas - need to make a note of this somewhere -->

@@

Part of the motivation for creating the `medrxivr` tool was a belief that the development and distribution of open source scripts and tools should be a fundamental part of evidence synthesis research.[@goldacre2019; @mckiernan2016] In the case of `medrxivr`, it is likely that several other evidence synthesists had written personal scripts that have a similar, or related, functionality - in fact, following development of the tool, I identified one other researcher that has done so (Nicholas Fraser, author of the `rbiorxiv` package, which allows for importing medRxiv metadata into R but does not provide search functionality).[@fraser2020rbiorixv] If these scripts continue to be developed in private and are never shared or publicised, this will inevitably hamper the efforts of evidence synthesis community, not only in terms of duplication of time and effort but also due to lost opportunities for collaboration.[@mckiernan2016] Creating and sharing well-documented packages, the recognised standard for sharing code in R, represents one way to reduce this inefficiency.[@vuorre2020]
